---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

```{r targets-setup, include=FALSE}
# library(targets)
# tar_option_set(envir = getNamespace("dftTrafficCounts"))
# tar_script()
# file.edit("_targets.R")
# tar_manifest()
```


# dftTrafficCounts

<!-- badges: start -->
[![R build status](https://github.com/itsleeds/dftTrafficCounts/workflows/R-CMD-check/badge.svg)](https://github.com/itsleeds/dftTrafficCounts/actions)
<!-- badges: end -->

The goal of dftTrafficCounts is to provide easy access to datasets provided by UK Department for Transport (DfT) on their [roadtraffic.dft.gov.uk](https://roadtraffic.dft.gov.uk) website.
See the [Metadata](http://data.dft.gov.uk.s3.amazonaws.com/road-traffic/all-traffic-data-metadata.pdf) document for details.

## Installation

<!-- You can install the released version of dftTrafficCounts from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("dftTrafficCounts") -->
<!-- ``` -->

<!-- And the development version from [GitHub](https://github.com/) with: -->

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("itsleeds/dftTrafficCounts")
```

<!-- From previous readme... -->
<!-- If you have traffic count data you would like to contribute to the Bristol City Council open data platform please contact city.transport@bristol.gov.uk -->

<!-- The entire dataset and a shapefile of the count locations can be downloaded here: http://1drv.ms/1LaIsZ4 -->

<!-- This data has not been quality checked and we strongly advise users conduct their own assessment of the quality of the data before using it. -->

Load the package (and packages we'll use below) as follows:

```{r}
library(dplyr)
library(ggplot2)
library(dftTrafficCounts)
```


# Getting traffic count data

You can read-in the data for a single counter point as follows:

```{r}
d_counter = dtc_import()
names(d_counter)
head(d_counter)
```

Read-in the data for major roads as follows:

```{r, eval=FALSE}
d_roads = dtc_import_roads()
names(d_roads)
```

Read-in data for all local authorities as follows:

```{r}
d_las = dtc_import_la()
head(d_las)
nrow(d_las)
length(unique(d_las$local_authority_id))
```

We can plot the results as follows:

```{r}
las_of_interest = c("Leeds", "Derby", "Southampton",
                    "Nottingham", "Birmingham")
d_las
d_las_of_interest = d_las %>%
   filter(local_authority_name %in% las_of_interest)
d_las %>% 
  ggplot() +
  geom_line(aes(year, all_motor_vehicles, group = local_authority_name), colour = "grey") +
  geom_line(aes(year, all_motor_vehicles, colour = local_authority_name), data = d_las_of_interest, size = 1.2)
```

```{r, eval=FALSE}
u = "http://data.dft.gov.uk/road-traffic/dft_traffic_counts_raw_counts.zip"
d = dtc_import(u = u)
```

```{r, echo=FALSE}
library(targets)
```


```{r, echo=FALSE}
d = tar_read(data)
```

This is a large dataset covering all local authorities in Great Britain:

```{r}
nrow(d)
ncol(d)
names(d)
head(d$local_authority_name)
```

```{r, eval=FALSE, echo=FALSE}
# aim: sense check
d_eastings_northings = d %>% 
  group_by(easting, northing) %>% 
  count()
d_sf = d_eastings_northings %>% sf::st_as_sf(coords = c("easting", "northing"), crs = 27700)
mapview::mapview(d_sf)
```


We can look at how patterns change over time for the whole dataset.

Lets look at the number of count entries for each year.

```{r, summary-count-day}
summary(d$count_date)
d_year_count = d %>% count(year)
d_day_count = d %>% count(count_date)
plot(d_day_count)
```


```{r summary-year}
# could become package function
tar_read(summary_mode) %>% 
  ggplot() +
  geom_line(aes(year, value, colour = name)) +
  scale_y_log10()
```

And by local authorities:

```{r lasinterest}
d_sample = d %>% filter(local_authority_name %in% las_of_interest)
d_summary_la = d_sample %>%
  select(pedal_cycles:buses_and_coaches | matches("year|local_authority_name")) %>% 
  group_by(year, local_authority_name) %>%
  summarise_all(sum) %>%
  tidyr::pivot_longer(cols = pedal_cycles:buses_and_coaches)
ggplot(d_summary_la) +
  geom_line(aes(year, value, colour = name)) +
  facet_wrap(~local_authority_name) +
  scale_y_log10()
```



# Reproducibility

You can reproduce the work presented here using the `targets` package:

```{r, echo=FALSE, message=FALSE, eval=FALSE}
remotes::install_github("wlandau/targets", quiet = TRUE)
remotes::install_github("wlandau/tarchetypes", quiet = TRUE)
```

```{r, eval=FALSE}
targets::tar_make()
```

Parts of the project are updated as follows:

```{r tarvis}
# tar_visnetwork()
```


# Summary of all the data

A summary of the raw DfT data is shown below:

```{r}
# print(tar_read(summary_dft))
```

